---
title: "Objective 1"
format: html
author: "Dewi"
---

# Load Library

```{r warning=FALSE}
# Data Input and Cleaning
library(readxl) 
library(openxlsx) 

library(ggmice) 

# Data Visualization
library(ggplot2)

# Clustering
library(mclust)  # Model-based clustering (Gaussian Mixture Models)
library(cluster) # General framework for various clustering algorithms

# Exploratory Data Analysis (EDA)
library(factoextra) # Perform EDA tasks (PCA, MCA, visualize categorical data)

# Additional Clustering Functions
library(fclust)  # Find optimal number of clusters, evaluate cluster validity, etc.

# Fuzzy c-means
library(e1071)

# Hierarchical Clustering Visualization
library(dendextend) # Enhance hierarchical clustering dendrogram visualizations


# Summarizing data
library(dplyr)
```

# Read and split data

```{r warning=FALSE}
data <- suppressMessages(read_xlsx(path = "data_master_project_3years.xlsx"))

# Split female and male into separate groups
male_data <- data[data$Sex == "Male",]
female_data <- data[data$Sex == "Female",]
```

# Data pre-processing

-   Calculating
-   Removing outliers (beyond 5 standard deviation)
-   Log transformation
-   Scaling

```{r}
preprocess_log_scale <- function(df) {
  # Replace spaces in column names with underscores
  colnames(df) <- gsub(" ", "_", colnames(df))
  
  # Check original column names
  #print("Original column names:")
  #print(colnames(df))
  
  # Convert the date column to Date type if it exists
  date_col_name <- "Date_E11_first_reported_(non-insulin-dependent_diabetes_mellitus)"
  if (date_col_name %in% colnames(df)) {
    print("Date column found. Converting to Date class...")
    df[[date_col_name]] <- as.Date(df[[date_col_name]])
  } else {
    stop("Date column not found in the data frame")
  }
  
  # Select clustering variables, dynamically include the identified date column
  subset_data <- df[, c(1, which(colnames(df) == "Month_of_birth"), 
                        which(colnames(df) == "Year_of_birth"), 
                        which(colnames(df) == date_col_name),
                        67, 7, 222, 234, 238, 256, 218)]
  
  # Rename columns for easier reference
  colnames(subset_data) <- c("ID", "Month_of_birth", "Year_of_birth", "Date_diagnosed", "BMI", "HbA1c", "Cholesterol", "HDL", "LDL", "Triglycerides", "CRP")
  
  # Check renamed column names
  print("Renamed column names:")
  print(colnames(subset_data))
  
  # Ensure month names are capitalized properly for matching
  subset_data$Month_of_birth <- tools::toTitleCase(tolower(subset_data$Month_of_birth))
  
  # Convert month and year of birth to numeric
  month_mapping <- c("January"=1, "February"=2, "March"=3, "April"=4, "May"=5, "June"=6,
                     "July"=7, "August"=8, "September"=9, "October"=10, "November"=11, "December"=12)
  subset_data$Month_of_birth <- as.numeric(month_mapping[subset_data$Month_of_birth])
  subset_data$Year_of_birth <- as.numeric(subset_data$Year_of_birth)
  
  # Calculate age at diagnosis
  diagnosis_year <- as.numeric(format(subset_data$Date_diagnosed, "%Y"))
  diagnosis_month <- as.numeric(format(subset_data$Date_diagnosed, "%m"))
  subset_data$age_at_diagnosis <- diagnosis_year - subset_data$Year_of_birth - 
                                  ifelse(diagnosis_month < subset_data$Month_of_birth, 1, 0)
  
  # Create numeric variables
  subset_data[, 5:12] <- apply(subset_data[, 5:12], 2, as.numeric)
  
  # Add Insulin resistance value
  subset <- data.frame(subset_data)
  subset$IR <- subset[, 10] / subset[, 8]
  subset <- subset[, -10]
  
  subset2 <- subset[complete.cases(subset), ]
  
  # Remove outliers
  for (col_name in colnames(subset2[, 5:12])) {
    col <- subset2[, col_name]
    mean_val <- mean(col)
    std_dev <- sd(col)
    subset2 <- subset2[!((col > mean_val + 5 * std_dev) | (col < mean_val - 5 * std_dev)), ]
  }
  
  subset2[,5:12] <- apply(subset2[,5:12], 2, log1p)
  subset2[,5:12] <- apply(subset2[,5:12], 2, scale)


  return(subset2)
}

female_processed <- preprocess_log_scale(female_data)
male_processed <- preprocess_log_scale(male_data)
```

check if data is processed well

```{r summary_processed}
summary(female_processed)
summary(male_processed)
```

-   Same function but without log transformation and scaling

```{r}
preprocess_outliers <- function(df) {
  # Replace spaces in column names with underscores
  colnames(df) <- gsub(" ", "_", colnames(df))
  
  # Check original column names
  #print("Original column names:")
  #print(colnames(df))
  
  # Convert the date column to Date type if it exists
  date_col_name <- "Date_E11_first_reported_(non-insulin-dependent_diabetes_mellitus)"
  if (date_col_name %in% colnames(df)) {
    print("Date column found. Converting to Date class...")
    df[[date_col_name]] <- as.Date(df[[date_col_name]])
  } else {
    stop("Date column not found in the data frame")
  }
  
  # Select clustering variables, dynamically include the identified date column
  subset_data <- df[, c(1, which(colnames(df) == "Month_of_birth"), 
                        which(colnames(df) == "Year_of_birth"), 
                        which(colnames(df) == date_col_name),
                        67, 7, 222, 234, 238, 256, 218)]
  
  # Rename columns for easier reference
  colnames(subset_data) <- c("ID", "Month_of_birth", "Year_of_birth", "Date_diagnosed", "BMI", "HbA1c", "Cholesterol", "HDL", "LDL", "Triglycerides", "CRP")
  
  # Check renamed column names
  print("Renamed column names:")
  print(colnames(subset_data))
  
  # Ensure month names are capitalized properly for matching
  subset_data$Month_of_birth <- tools::toTitleCase(tolower(subset_data$Month_of_birth))
  
  # Convert month and year of birth to numeric
  month_mapping <- c("January"=1, "February"=2, "March"=3, "April"=4, "May"=5, "June"=6,
                     "July"=7, "August"=8, "September"=9, "October"=10, "November"=11, "December"=12)
  subset_data$Month_of_birth <- as.numeric(month_mapping[subset_data$Month_of_birth])
  subset_data$Year_of_birth <- as.numeric(subset_data$Year_of_birth)
  
  # Calculate age at diagnosis
  diagnosis_year <- as.numeric(format(subset_data$Date_diagnosed, "%Y"))
  diagnosis_month <- as.numeric(format(subset_data$Date_diagnosed, "%m"))
  subset_data$age_at_diagnosis <- diagnosis_year - subset_data$Year_of_birth - 
                                  ifelse(diagnosis_month < subset_data$Month_of_birth, 1, 0)
  
  # Create numeric variables
  subset_data[, 5:12] <- apply(subset_data[, 5:12], 2, as.numeric)
  
  # Add Insulin resistance value
  subset <- data.frame(subset_data)
  subset$IR <- subset[, 10] / subset[, 8]
  subset <- subset[, -10]
  
  subset2 <- subset[complete.cases(subset), ]
  
  # Remove outliers
  for (col_name in colnames(subset2[, 5:12])) {
    col <- subset2[, col_name]
    mean_val <- mean(col)
    std_dev <- sd(col)
    subset2 <- subset2[!((col > mean_val + 5 * std_dev) | (col < mean_val - 5 * std_dev)), ]
  }
  

  return(subset2)
}

female_processed_no_scale_log <- preprocess_outliers(female_data)
male_processed_no_scale_log <- preprocess_outliers(male_data)
```

```{r hist}
# Define the range of columns
columns_range <- 5:12

# Create a new plotting window
par(mfrow=c(2, 4)) 

# Loop through each column and plot histogram
for (col_index in columns_range) {
  # Get column name
  col_name <- names(male_processed_no_scale_log)[col_index]
  
  # Plot histogram
  hist(male_processed_no_scale_log[, col_index], main = paste("Histogram of Male", col_name), xlab = col_name, col = "skyblue")
}

# Create a new plotting window
par(mfrow=c(2, 4)) 

# Loop through each column and plot histogram
for (col_index in columns_range) {
  # Get column name
  col_name <- names(male_processed)[col_index]
  
  # Plot histogram
  hist(male_processed[, col_index], main = paste("Histogram of Male", col_name), xlab = col_name, col = "skyblue")
}

# Reset plotting layout
par(mfrow=c(1, 1))
```

# Clustering

## Optimal clustering number

```{r}
set.seed(123)
```

Cluster tendency with Hopkins statistics

```{r eval=FALSE}
# Compute Hopkins statistic 
result_tendency_female <- get_clust_tendency(female_processed[, 5:12], n = nrow(female_processed)-1, graph = FALSE)
result_tendency_female$hopkins_stat
```

A hopkins statistics of 0.77, indicate strong cluster tendency. The data likely has well-defined clusters.

Finding optimal cluster number with Sum of Squares, Silhouette score and GAP-distance for females:

```{r optimal_female}
fviz_nbclust(female_processed[, 5:12], kmeans, method = "wss")
fviz_nbclust(female_processed[, 5:12], kmeans, method = "silhouette")
fviz_nbclust(female_processed[, 5:12], kmeans, method = "gap_stat", iter.max=50)
```
Finding optimal cluster number with Sum of Squares, Silhouette score and GAP-distance for males:
```{r optimal_male}
fviz_nbclust(male_processed[, 5:12], kmeans, method = "wss", iter.max=50)
fviz_nbclust(male_processed[, 5:12], kmeans, method = "silhouette")
fviz_nbclust(male_processed[, 5:12], kmeans, method = "gap_stat", iter.max=50)
```

3 clusters seems like the best amount for both female and male.

We can perform a cluster stability test.
```{r}
library(fpc)

# Perform cluster-wise stability evaluation using clusterboot
cluster_stability <- clusterboot(data = female_processed[, 5:12], B=1000, clustermethod=kmeansCBI, k=3, count=FALSE )

# AvgJaccard <0.6 is unstable & >0.85 is highly stable
AvgJaccard <- cluster_stability$bootmean
Instability <- cluster_stability$bootbrd/1000
Clusters <- c(1:4)
Eval <- cbind(Clusters, AvgJaccard, Instability)
Eval 
```
This test gives a stable result for k-means with 3 clusters.

To visualize the clusters in 2 dimensions, PCA can be performed. PCA reduces the 8 variables into 1 dimension. 

PCA k-means clustering Female 4 clusters:

```{r}
fviz_cluster(kmeans(female_processed[,5:12], centers = 4, iter.max = 100, nstart = 100), data= female_processed[, 5:12])
```

PCA k-means clustering Female 3 clusters:

```{r}
fviz_cluster(kmeans(female_processed[,5:12], centers = 3, iter.max = 100, nstart = 100), data= female_processed[, 5:12])
```

-   Dim1 (Principal Component 1): This component captures the majority of the variance in your dataset, specifically 27.2%. It suggests that the patterns or structures present in your data are primarily aligned with this component. In other words, "dim1" summarizes most of the important information contained in your original dataset.
-   Dim2 (Principal Component 2): This component explains a smaller proportion of the total variance, specifically 26.1%. While it captures less variability compared to "dim1", it still contains some valuable information about the dataset, albeit to a lesser extent.

PCA k-means clustering Male 4 clusters:

```{r}
fviz_cluster(kmeans(male_processed[,5:12], centers = 4, iter.max = 100, nstart = 100), data= male_processed[, 5:12])
```

PCA k-means clustering Male 3 clusters:

```{r}
fviz_cluster(kmeans(male_processed[,5:12], centers = 3, iter.max = 100, nstart = 100), data= male_processed[, 5:12])
```

4 clusters does indeed not look like the proper amount of clusters, as there is overlap in the clusters. This confirms that 3 clusters is the best custer amount.

-   Dim1 (Principal Component 1): This component captures the majority of the variance in your dataset, specifically 26.6%. It suggests that the patterns or structures present in your data are primarily aligned with this component. In other words, "dim1" summarizes most of the important information contained in your original dataset.
-   Dim2 (Principal Component 2): This component explains a smaller proportion of the total variance, specifically 25.1%. While it captures less variability compared to "dim1", it still contains some valuable information about the dataset, albeit to a lesser extent.

WSS and the silhouette scores for 3 clusters for the female and male datasets will be used to evaluate the internal clustering performance. The WSS provides a measure of compactness, while the silhouette score provides insight into the separation between clusters. Both metrics together can give you a comprehensive view of the clustering performance.


The Within Sum of Squares (WSS), also known as the Within-Cluster Sum of Squares, is a measure of the compactness of clusters in clustering analysis. It quantifies the variability of the data points within each cluster and is a key metric in assessing the performance of clustering algorithms like K-means.

WSS Definition
WSS is the sum of squared distances between each data point and the centroid of the cluster it belongs to. It can be calculated both for the entire clustering solution (total WSS) and individually for each cluster.

* Total WSS: Sum of the WSS values for all clusters.
* Cluster-wise WSS: WSS value for each individual cluster.


## Hard Clustering

### K-means

```{r}
# setting the cluster amount to 3 clusters
cl = 3
```

```{r}
# K-means on dataset with 3 clusters
kmeans_f <-  kmeans(female_processed[, 5:12], centers = cl, nstart = 100, iter.max = 100000)
kmeans_m <- kmeans(male_processed[,5:12], centers = cl, nstart = 100, iter.max = 100000)
```

#### K-means female results

```{r}
# Adding k-means clusters to data frames
cluster_km_f <- data.frame(kmeans_f$cluster)
data_f <- cbind(female_processed, cluster_km_f)
female_processed_no_scale_log <- cbind(female_processed_no_scale_log, cluster_km_f)

# Boxplot
for (col in names(female_processed_no_scale_log[,5:12])) {
  boxplot(female_processed_no_scale_log[,col] ~ female_processed_no_scale_log[["kmeans_f.cluster"]], 
        xlab = "cluster",
        ylab = col,
        main = col)
}
# the mean of each variable by clusters
aggregate(female_processed_no_scale_log[,5:12], by = list(cluster = kmeans_f$cluster), mean)

# Calculate silhouette scores per cluster
silhouette_scoresf <- silhouette(kmeans_f$cluster, dist(female_processed[,5:12]))
summary(silhouette_scoresf)
```

```{r}
# Calculate WSS for each cluster and total WSS for female dataset
wss_clusters_f <- kmeans_f$withinss
total_wss_f <- sum(wss_clusters_f)
cat("Within-cluster Sum of Squares (WSS) for female dataset with 3 clusters:\n")
cat("Cluster-wise WSS:", wss_clusters_f, "\n")
cat("Total WSS:", total_wss_f, "\n")

```


#### K-means male results

```{r}
# Adding k-means clusters to data frames
cluster_km_m <- data.frame(kmeans_m$cluster)
data_m <- cbind(male_processed, cluster_km_m)
male_processed_no_scale_log <- cbind(male_processed_no_scale_log, cluster_km_m)

# Boxplot
for (col in names(male_processed_no_scale_log[,5:12])) {
  boxplot(male_processed_no_scale_log[,col] ~ male_processed_no_scale_log[["kmeans_m.cluster"]], 
        xlab = "cluster",
        ylab = col,
        main = col)
}
# the mean of each variable by clusters
aggregate(male_processed_no_scale_log[,5:12], by = list(cluster = kmeans_m$cluster), mean)

# Calculate silhouette scores per cluster
silhouette_scoresm <- silhouette(kmeans_m$cluster, dist(male_processed[,5:12]))

summary(silhouette_scoresm)

# Calculate WSS for each cluster and total WSS for male dataset
wss_clusters_m <- kmeans_m$withinss
total_wss_m <- sum(wss_clusters_m)
cat("Within-cluster Sum of Squares (WSS) for male dataset with 3 clusters:\n")
cat("Cluster-wise WSS:", wss_clusters_m, "\n")
cat("Total WSS:", total_wss_m, "\n")
```

The preliminary outcomes of k-means appear promising, with the silhouette scores being satisfactory.


## Soft Clustering

### Fuzzy c-means

m = 2: This is the fuzziness parameter, also known as the fuzzifier or exponent. It determines the degree of fuzziness in the clustering. When m = 1, the algorithm becomes equivalent to hard k-means clustering. As m increases, the clusters become fuzzier. A common choice for this parameter is 2.

#### fuzzy c-means female results

```{r}
# perform fuzzy k means (soft clustering)
fuzzy_cmeans_female <- cmeans(female_processed[, 5:12], centers = cl, m = 2, iter.max = 100)
#summary(fuzzy_cmeans_female)
fuzzy_cluster_f <- fuzzy_cmeans_female$cluster

# Adding fuzzy c-means clusters to data frames
data_f <- cbind(data_f, fuzzy_f = fuzzy_cluster_f)
 female_processed_no_scale_log <- cbind(female_processed_no_scale_log, fuzzy_f = fuzzy_cluster_f)


# Boxplot
for (col in names(female_processed_no_scale_log[,5:12])) {
  boxplot(female_processed_no_scale_log[[col]] ~ female_processed_no_scale_log[["fuzzy_f"]], 
        xlab = "cluster",
        ylab = col,
        main = col)
}
# the mean of each variable by clusters
aggregate(female_processed_no_scale_log[,5:12], by = list(cluster = fuzzy_cluster_f), mean)

# cluster size
fuzzy_cmeans_female$size

fuzzy_silh_scores_f <- silhouette(fuzzy_cluster_f, dist(female_processed[,5:12]))

summary(fuzzy_silh_scores_f)
```
```{r}
# Calculate WSS for each cluster
wss_fuzzy <- function(data, centers, membership) {
  wss_values <- numeric(nrow(centers))
  for (k in 1:nrow(centers)) {
    # Calculate the weighted squared distances for cluster k
    dists <- (data - centers[k, ])^2
    wss_values[k] <- sum(membership[, k] * rowSums(dists))
  }
  return(wss_values)
}
# Extract centers and membership values from fuzzy c-means result
centers_f <- fuzzy_cmeans_female$centers
membership_f <- fuzzy_cmeans_female$membership

# Calculate WSS for each cluster
wss_clusters_f <- wss_fuzzy(female_processed[, 5:12], centers_f, membership_f)
total_wss_f <- sum(wss_clusters_f)

# Print WSS values
cat("Within-cluster Sum of Squares (WSS) for female dataset with fuzzy c-means clustering:\n")
cat("Cluster-wise WSS:", wss_clusters_f, "\n")
cat("Total WSS:", total_wss_f, "\n")

```

#### fuzzy c-means male results

```{r}
# perform fuzzy k means (soft clustering)
fuzzy_cmeans_male <- cmeans(male_processed[, 5:12], centers = cl, m = 2, iter.max = 100)
#summary(fuzzy_cmeans_male)
fuzzy_cluster_m <- fuzzy_cmeans_male$cluster

# Adding gmm clusters to data frames
data_m <- cbind(data_m, fuzzy_m = fuzzy_cluster_m)
male_processed_no_scale_log <- cbind(male_processed_no_scale_log, fuzzy_m = fuzzy_cluster_m)


# Boxplot
for (col in names(male_processed_no_scale_log[,5:12])) {
  boxplot(male_processed_no_scale_log[[col]] ~ male_processed_no_scale_log[["fuzzy_m"]], 
        xlab = "cluster",
        ylab = col,
        main = col)
}
# the mean of each variable by clusters
aggregate(male_processed_no_scale_log[,5:12], by = list(cluster = fuzzy_cluster_m), mean)

# cluster size
fuzzy_cmeans_male$size

fuzzy_silh_scores_m <- silhouette(fuzzy_cluster_m, dist(male_processed[,5:12]))

summary(fuzzy_silh_scores_m)


```
```{r}
# Extract centers and membership values from fuzzy c-means result
centers_m <- fuzzy_cmeans_male$centers
membership_m <- fuzzy_cmeans_male$membership

# Calculate WSS for each cluster
wss_clusters_m <- wss_fuzzy(male_processed[, 5:12], centers_m, membership_m)
total_wss_m <- sum(wss_clusters_m)

# Print WSS values
cat("Within-cluster Sum of Squares (WSS) for male dataset with fuzzy c-means clustering:\n")
cat("Cluster-wise WSS:", wss_clusters_m, "\n")
cat("Total WSS:", total_wss_m, "\n")

```


# Export data

```{r}
female_xlsx_mean_k3 <- aggregate(female_processed_no_scale_log[,5:12], by = list(cluster = kmeans_f$cluster), mean)

# naming clusters based on averages

# Function to rename clusters based on conditions
rename_clusters_female <- function(cluster_means) {
  renamed_clusters <- character(nrow(cluster_means))
  
  for (i in 1:nrow(cluster_means)) {
    if (cluster_means$BMI[i] > 36.5) {
      renamed_clusters[i] <- "Obesity-related"
    } else if (cluster_means$age_at_diagnosis[i] > 61) {
      renamed_clusters[i] <- "Age-related"
    } else if (cluster_means$Cholesterol[i] > 6) {
      renamed_clusters[i] <- "Lipid-related"
    } else {
      renamed_clusters[i] <- paste("Cluster", cluster_means$cluster[i])
    }
  }
  
  return(renamed_clusters)
}

# Rename clusters for female dataset
renamed_clusters_f <- rename_clusters_female(female_xlsx_mean_k3)

# Map renamed clusters back to the data for female dataset
female_processed_no_scale_log$cluster <- renamed_clusters_f[kmeans_f$cluster]

##########################
male_xlsx_mean_k3 <- aggregate(male_processed_no_scale_log[,5:12], by = list(cluster = kmeans_m$cluster), mean)

# naming clusters based on averages

# Function to rename clusters based on conditions
rename_clusters_male <- function(cluster_means) {
  renamed_clusters <- character(nrow(cluster_means))
  
  for (i in 1:nrow(cluster_means)) {
    if (cluster_means$BMI[i] > 32.8) {
      renamed_clusters[i] <- "Obesity-related"
    } else if (cluster_means$age_at_diagnosis[i] > 62) {
      renamed_clusters[i] <- "Age-related"
    } else if (cluster_means$Cholesterol[i] > 5.5) {
      renamed_clusters[i] <- "Lipid-related"
    } else {
      renamed_clusters[i] <- paste("Cluster", cluster_means$cluster[i])
    }
  }
  
  return(renamed_clusters)
}

# Rename clusters for male dataset
renamed_clusters_m <- rename_clusters_male(male_xlsx_mean_k3)

# Map renamed clusters back to the data for female dataset
male_processed_no_scale_log$cluster <- renamed_clusters_m[kmeans_m$cluster]
```


```{r eval = FALSE}
# write data for females and males to xlsx file with clusters


# write aggregated data to xlsx

write.xlsx(female_xlsx_mean_k3, file = "kmeans_female_mean_k3.xlsx")
write.xlsx(male_xlsx_mean_k3, file = "kmeans_male_mean_k3.xlsx")

# write aggregated data to xlsx
female_xlsx_mean_c3 <- aggregate(female_processed_no_scale_log[,5:12], by = list(cluster = fuzzy_cmeans_female$cluster), mean)
write.xlsx(female_xlsx_mean_c3, file = "cmeans_female_mean_c3.xlsx")

male_xlsx_mean_c3 <- aggregate(male_processed_no_scale_log[,5:12], by = list(cluster = fuzzy_cmeans_male$cluster), mean)
write.xlsx(male_xlsx_mean_c3, file = "cmeans_male_mean_c3.xlsx")

# Remove the last column from the data frames
female_final <- female_processed_no_scale_log[, -c(ncol(female_processed_no_scale_log) - 1, ncol(female_processed_no_scale_log) - 2)]
male_final <- male_processed_no_scale_log[, -c(ncol(male_processed_no_scale_log) - 1, ncol(male_processed_no_scale_log) - 2)]

write.xlsx(female_final, file = "female_final.xlsx", rowNames = FALSE)
write.xlsx(male_final, file = "male_final.xlsx", rowNames = FALSE)
```

# Summarize data

```{r}
# Summarize min and max per variable per cluster
summary_stats_km_f <- female_processed_no_scale_log %>%
  group_by(cluster) %>%
  summarize(across(5:12, list(min = min, max = max), .names = "{.col}_{.fn}"))

print(summary_stats_km_f)
```

```{r}
# Summarize min and max per variable per cluster
summary_stats_cm_f <- female_processed_no_scale_log %>%
  group_by(fuzzy_f) %>%
  summarize(across(5:12, list(min = min, max = max), .names = "{.col}_{.fn}"))

print(summary_stats_cm_f)
```

```{r}
# Summarize min and max per variable per cluster
summary_stats_km_m <- male_processed_no_scale_log %>%
  group_by(cluster) %>%
  summarize(across(5:12, list(min = min, max = max), .names = "{.col}_{.fn}"))

print(summary_stats_km_m)
```

```{r}
# Summarize min and max per variable per cluster
summary_stats_cm_m <- male_processed_no_scale_log %>%
  group_by(fuzzy_m) %>%
  summarize(across(5:12, list(min = min, max = max), .names = "{.col}_{.fn}"))

print(summary_stats_cm_m)
```

# Visualise Data with PCA

```{r}
# Perform PCA on the relevant columns (columns 5 to 12)
pca_f <- prcomp(female_processed[, 5:12], scale. = TRUE)

# Get the PCA results and convert to a data frame
pca_f_df <- as.data.frame(pca_f$x)

# Add cluster assignments to the PCA results
pca_f_df$kmeans <- kmeans_f$cluster

pca_f_df$cmeans <- fuzzy_cmeans_female$cluster

```

```{r}
# Perform PCA on the relevant columns (columns 5 to 12)
pca_m <- prcomp(male_processed[, 5:12], scale. = TRUE)

# Get the PCA results and convert to a data frame
pca_m_df <- as.data.frame(pca_m$x)

# Add cluster assignments to the PCA results
pca_m_df$kmeans <- kmeans_m$cluster

pca_m_df$cmeans <- fuzzy_cmeans_male$cluster

```

## PCA k-means visualisation

```{r}
# Plotting the first two principal components
ggplot(pca_f_df, aes(x = PC1, y = PC2, color = factor(kmeans))) +
  geom_point(size = 2) +
  labs(title = "PCA of Female Data with kmeans clusters",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Cluster") +
  theme_minimal()
```

```{r}
# Plotting the first two principal components
ggplot(pca_f_df, aes(x = PC1, y = PC2, color = factor(cmeans))) +
  geom_point(size = 2) +
  labs(title = "PCA of Female Data with c-means clusters",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Cluster") +
  theme_minimal()
```

```{r}
# Plotting the first two principal components
ggplot(pca_m_df, aes(x = PC1, y = PC2, color = factor(kmeans))) +
  geom_point(size = 2) +
  labs(title = "PCA of Male Data with kmeans clusters",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Cluster") +
  theme_minimal()
```

```{r}
# Plotting the first two principal components
ggplot(pca_m_df, aes(x = PC1, y = PC2, color = factor(cmeans))) +
  geom_point(size = 2) +
  labs(title = "PCA of Male Data with c-means clusters",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Cluster") +
  theme_minimal()
```
